---
title: "VST kaplan mier curve"
author: "Tim N"
date: "9/15/2021"
output: html_document
---

The goal of this document is to make a kaplan mier curve for each cancer using
the top 10 genes and a respective Z-score.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load librarries
```{r}
library(tidyverse)
library(DESeq2)
library(recount3)
library(survival)
library(ggfortify)
library(patchwork)
```


```{r}
genes_of_int_high <- c("COL10A1", "MMP11", "COL11A1", "COMP", "SPP1", "MMP9", "MMP1", "CST1", "CTHRC1", "MMP7")
genes_of_int_low <- c("MASP1", "PCOLCE2", "SRPX", "WISP2", "LGI4", "MMRN1", "OGN", "TNXB", "DPT", "ADAM33")
cancers <- c("BRCA", "LUAD", "COAD", "THCA", "PRAD", "ESCA", "OV", "STAD", "PAAD", "LIHC")
```
Load the TCGA project
```{r}
human_projects <- available_projects()
tcga_projects <- subset(human_projects, file_source == "tcga" & project %in% cancers)

```
Make projects in list by looping through and acquiring the data
```{r}
rse_list <- list()
for (cur_cancer in cancers) {
    cur_proj <- subset(tcga_projects, project == cur_cancer)
    cur_rse <- create_rse(cur_proj)
    rse_list[[cur_cancer]] <- cur_rse
}
```

This is the main loop, we will be using this to begin our analysis of the RNA-seq data
and save the results
```{r}
## To do: Use the upper and lower quantile of individuals
## To do: Make sure people who survive stay in
result_list <- list()
result_table <- NULL
aova_df_out <-  NULL
linear_model_df_out <- NULL
for (tcga_ind in seq_along(rse_list)) {
    ### 1. Pull out the data required
    ## Pull out the first cancer type
    cur_rse <- rse_list[[tcga_ind]]
    ## tcga cancer name
    tcga_cancer_name <- names(rse_list)[tcga_ind]
    ## Filter out normal and metastic
    cur_coldat <- colData(cur_rse)
    cancer_ind <- cur_coldat$tcga.cgc_sample_sample_type %in% "Primary Tumor"
    cur_counts <- assay(cur_rse)[,cancer_ind]
    ### 2. Normalize the data using a vst
    cur_vst <- vst(cur_counts)
    ## Pull row data
    cur_row_dat <- rowRanges(cur_rse)
    ### 3. Filter the data to genes we care about
    ## High genes
    high_ind <- (cur_row_dat$gene_name %in% genes_of_int_high &
                     startsWith(as.character(as.data.frame(cur_row_dat)$seqnames), "chr"))
    high_rownames <- rowRanges(cur_rse)[high_ind,]
    high_vst <- cur_vst[high_ind,]
    if (identical(names(high_rownames), row.names(high_vst)) == FALSE) {
        stop()
    }
    row.names(high_vst) <- high_rownames$gene_name
    ## Low genes
    low_ind <- (cur_row_dat$gene_name %in% genes_of_int_low &
                     startsWith(as.character(as.data.frame(cur_row_dat)$seqnames), "chr"))
    low_rownames <- rowRanges(cur_rse)[low_ind,]
    low_vst <- cur_vst[low_ind,]
    if (identical(names(low_rownames), row.names(low_vst)) == FALSE) {
        stop()
    }
    row.names(low_vst) <- low_rownames$gene_name
    ## put them into list
    h_l_list <- list()
    h_l_list[["high"]] <- high_vst
    h_l_list[["low"]] <- low_vst
    z_score_out <- NULL
    ### 4. Create two Z-scores for the high gene list and low
    for (vst_ind in seq_along(h_l_list)) {
        hl_vst <- h_l_list[[vst_ind]]
        hl_vst_centered <- sweep(hl_vst, MARGIN = 1, STATS = rowMeans(hl_vst), FUN = "-")
        hl_vst_centered_standard <- sweep(hl_vst_centered, MARGIN = 1,
                                          STATS = sd(hl_vst_centered), FUN = "/")
        z_score_out <- cbind(z_score_out, colMeans(hl_vst_centered_standard))
    }
    colnames(z_score_out) <- paste0(names(h_l_list), sep = "_zscore")
    ### 5. Create a Kaplan mier curve result and plot and store it in a nested
    ###     list.
    filt_coldat <- as.data.frame(cur_coldat[cancer_ind,])
    filt_coldat <- select(filt_coldat, starts_with("tcga.cgc")) %>%
        rownames_to_column(var = "samp_id")
    z_score_out <- as.data.frame(z_score_out) %>%
        rownames_to_column(var = "samp_id")
    all_dat_pre_coding <- left_join(z_score_out, filt_coldat) %>%
        #filter(tcga.cgc_case_days_to_death != 0) %>%
        mutate(high_group = ifelse(high_zscore > median(high_zscore),
                                                        "high_exprs",
                                                        "low_exprs"),
               ## Make high and low quantiles
               high_quant_split = case_when(
                   high_zscore >= quantile(high_zscore, .75) ~ "High Z-score Exprs",
                   high_zscore <= quantile(high_zscore, .25) ~ "Low Z-score Exprs",
                   TRUE ~ "middle_exprs"
               ),
               low_quant_split = case_when(
                   low_zscore >= quantile(low_zscore, .75) ~ "High Z-score Exprs",
                   low_zscore <= quantile(low_zscore, .25) ~ "Low Z-score Exprs",
                   TRUE ~ "middle_exprs"
               ),
               ## Low median group
               low_group = ifelse(low_zscore > median(low_zscore),
                                                        "high_exprs",
                                                        "low_exprs")
        )
    ## Need to carefully change the data to fit the analysis
    all_dat <- all_dat_pre_coding %>%
        mutate(
               ## Make status numweric
               status = case_when(
                   tcga.cgc_follow_up_vital_status == "Alive" ~ 0,
                   tcga.cgc_follow_up_vital_status == "Dead" ~ 1,
                   is.numeric(tcga.cgc_case_days_to_death) & is.na(tcga.cgc_follow_up_vital_status) ~ 1
               ),
               # status = ifelse(tcga.cgc_follow_up_vital_status == "Alive",
               #                 0, 1),
               ## Change days to death to drop 0s
               days_to_death = ifelse(is.na(tcga.cgc_case_days_to_death),
                                      ## Make follow up the censor date
                                     tcga.cgc_case_days_to_last_follow_up,
                                     tcga.cgc_case_days_to_death)) 
    
    
    high_dat <- filter(all_dat, high_quant_split != "middle_exprs") 
    low_dat <- filter(all_dat, low_quant_split != "middle_exprs")
    
    high_fit <- survfit(Surv(high_dat$days_to_death,
                             event = high_dat$status) ~ high_group, data = high_dat)
    low_fit <- survfit(Surv(days_to_death,
                            status) ~ low_group, data = low_dat)
    high_plt <- survminer::ggsurvplot(high_fit) +
        labs(title = "Genes that increase in cancer Z-score survival",
             subtitle = paste0(survminer::surv_pvalue(high_fit)$pval.txt,
                               "; N high_exprs = ", high_fit$n[1],
                               "; N low_exprs = ", high_fit$n[2]))
    if (sum(low_fit$n )> 1) {
       low_plt <- survminer::ggsurvplot(low_fit) +
        labs(title = "Genes that decrease in cancer Z-score survival",
             subtitle = paste0(survminer::surv_pvalue(low_fit)$pval.txt,
                               "; N high_exprs = ", low_fit$n[1],
                               "; N low_exprs = ", low_fit$n[2])) 
    } else{
        low_plt <- ggplot()
    }
    
    plot_list <- list()
    plot_list[["high_plot"]] <- high_plt
    plot_list[["low_plot"]] <- low_plt
    result_list[[]] <- plot_list
    
    ## Table of results
    high_surv_medians <- survminer::surv_median(high_fit)[,1:2]
    ## Catch if there aren't enough values for a median 
    if (sum(is.na(high_surv_medians$median)) > 0) {
        high_higher_curve = NA
    } else{
        if (high_surv_medians$median[1] > high_surv_medians$median[2]) {
            high_higher_curve = str_split(high_surv_medians$strata[1], "=")[[1]][2]
        } else{
            high_higher_curve = str_split(high_surv_medians$strata[2], "=")[[1]][2]
        }
    }
    high_pval = survminer::surv_pvalue(high_fit)$pval
    high_sig = high_pval < 0.05
    row_high <- cbind(high_higher_curve, high_pval, high_sig, "rank up genes", )
    
    low_surv_medians <- survminer::surv_median(low_fit)[,1:2]
    ## Catch if there aren't enough values for a median
    if (sum(is.na(low_surv_medians$median)) > 0) {
        low_higher_curve = NA
    } else{
        if (low_surv_medians$median[1] > low_surv_medians$median[2]) {
            low_higher_curve = str_split(low_surv_medians$strata[1], "=")[[1]][2]
        } else{
            low_higher_curve = str_split(low_surv_medians$strata[2], "=")[[1]][2]
        }
    }
    low_pval = survminer::surv_pvalue(low_fit)$pval
    low_sig = low_pval < 0.05
    row_low <- cbind(low_higher_curve, low_pval, low_sig, "rank down genes", )
    new_rows <- rbind(row_high, row_low)
    result_table <- rbind(result_table, new_rows)
    
    ### Cancer stage bins
    # table(high_dat$tcga.cgc_case_pathologic_stage, high_dat$high_quant_split)
    # table(low_dat$tcga.cgc_case_pathologic_stage, low_dat$low_quant_split)
    # ggplot(high_dat, (aes(tcga.cgc_case_pathologic_stage, fill = high_quant_split))) +
    #     geom_bar(position = "dodge")
    # ggplot(low_dat, (aes(tcga.cgc_case_pathologic_stage, fill = low_quant_split))) +
    #     geom_bar(position = "dodge") +
    #     coord_flip()
    # ggplot(all_dat, (aes(tcga.cgc_case_pathologic_stage, low_zscore))) +
    #     ggforce::geom_sina() 
    # ggplot(all_dat, (aes(tcga.cgc_case_pathologic_stage, high_zscore))) +
    #     ggforce::geom_sina() 
    ### Cancer stage anovas
    high_anova <- broom::tidy(
        anova(aov(high_zscore ~ tcga.cgc_case_pathologic_stage,
                  data = all_dat)))[1,]
    low_anova <- broom::tidy(
        anova(aov(low_zscore ~ tcga.cgc_case_pathologic_stage,
                  data = all_dat)))[1,]
    new_rows_aov <- NULL
    new_rows_aov <- rbind(high_anova, low_anova)
    new_rows_aov$rank_genes <- c("increase_in_cancer", "decrease_in_cancer")
    new_rows_aov$tcga_cancer <-
    aov_df_out <- rbind(aova_df_out, new_rows_aov)
    ### Cancer stage ordinal linear models 
    cancer_factor_levels <- c("Stage I", "Stage II",
                              "Stage IIIA","Stage IIIB","Stage IIIC",
                              "Stage IV", "Stage IVA","Stage IVB")
    all_dat$pathologic_stage_factor <- factor(all_dat$tcga.cgc_case_pathologic_stage,
                                                     levels = cancer_factor_levels,
                                                     ordered = T)
    stage_upreg_model <- lm(high_zscore ~ pathologic_stage_factor,
                            data = all_dat)
    lm_row_upreg <- cbind(broom::tidy(summary(stage_upreg_model))[2,],
                    direction = "upregulated", tcga_cancer_name)
    stage_downreg_model <- lm(low_zscore ~ pathologic_stage_factor,
                            data = all_dat)
    lm_row_downreg <- cbind(broom::tidy(summary(stage_downreg_model))[2,],
                    direction ="downregulated", tcga_cancer_name)
    
    linear_model_df_out <- rbind(linear_model_df_out, lm_row_upreg, lm_row_downreg)
    
}
```

PLot out
```{r}
pdf("data_out/kaplan_z_scores.pdf",width = "12", height = "15")
for (tcga_ind in seq_along(result_list)) {
    ## Skip THCA
    # if (tcga_ind == 4) {
    #     next()
    # }
    cur_tcga <- names(result_list)[tcga_ind]
    high_plt <- result_list[[tcga_ind]][[1]]
    low_plt <- result_list[[tcga_ind]][[2]]
    final_plot <- high_plt$plot / low_plt$plot + plot_annotation(title = cur_tcga)
    print(final_plot)
}
dev.off()
```
Make a plot of the results

```{r}
results_df <- as.data.frame(result_table) %>%
    mutate(high_pval = as.numeric(high_pval),
           corrected_pval = p.adjust(high_pval,method = "fdr"),
           corrected_pval_sig = corrected_pval < 0.05) %>%
    dplyr::rename(greater_median = high_higher_curve,
                  pvalue = high_pval,
                  sig_pval = high_sig,
                  rank_change_direction = V4,
                  tcga_cancer = V5)
sig_results <- filter(results_df, corrected_pval_sig == TRUE)

table(sig_results$tcga_cancer, sig_results$rank_change_direction)
table(sig_results$greater_median, sig_results$rank_change_direction)
```




Sanity checks
```{r}
mmp11_ind <- (rowRanges(cur_rse)$gene_name %in% "MMP11")
rowRanges(cur_rse)[mmp11_ind,]


mmp_dat <- as.data.frame(t(cur_vst[mmp11_ind,]))

plot(mmp_dat$ENSG00000275365.3, mmp_dat$ENSG00000099953.9)

```

